---
title: "p8105_hw6_zw2974"
author: "Zihan Wu"
date: "2023-11-18"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
library(boot)
library(readr)
library(purrr)
library(mgcv)
library(modelr)
set.seed(1)
```

## Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Problem 2
#### Download data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

#### Bootstrap Analysis
```{r}
boot_sample = function(weather_df) {
  sample_n(weather_df, size = nrow(weather_df), replace = TRUE)
}
```

#### 5000 bootstrap samples
```{r}
boot_strap = weather_df |>  
  modelr::bootstrap(n = 5000) |> 
  mutate(
    mods = map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
    mods_beta = map(mods, broom::glance),
    mods_r2 = map(mods, broom::tidy)) |> 
  unnest(mods_r2, mods_beta) |> 
  dplyr::select(.id, term, estimate, r.squared) |> 
  pivot_wider(names_from = term, values_from = estimate) |> 
  rename(
    beta_0 = `(Intercept)`,
    beta_1 = tmin,
    beta_2 = prcp) 

res = boot_strap |> 
  summarise(r.squared = r.squared,
    log_beta12 = log(beta_1 * beta_2),
    log_beta01 = log(beta_0 * beta_1))
```

#### CI for r-squared and log
```{r}
# calculate CI for r-squared and log
r2_ci = 
  quantile(pull(res,r.squared), probs = c(0.025,0.975)) |> 
  knitr::kable()

log_ci = 
  quantile(pull(res,log_beta01), probs = c(0.025,0.975)) |>  
  knitr::kable()
  
r2_ci
log_ci
```
#### Distribution Plots
```{r}
r2_dist = res |> 
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(title = "Distribution of Bootstrapped R-squared Estimates", x = "R-squared", y = "Density")
r2_dist
```

This plot is a density plot of bootstrapped R-squared estimates. The distribution of the R-squared values is unimodal and shows a peak around 0.92, indicating that the majority of the bootstrapped samples have an R-squared value close to this number. Also, the distribution is slightly skewed to the left, but is approximately as normal distribution.

```{r}
b1b2_dist = res |> 
  ggplot(aes(x = log_beta12)) +
  geom_density() +
  labs(title = "Distribution of Bootstrapped log(beta1*beta2) Estimates", x = "log(beta1*beta2)", y = "Density")
b1b2_dist
```

This plot is a density plot of bootstrapped Log Beta Product estimates. The distribution appears to be left-skewed, which may be caused by the `NA` generated during the logarithm calculation.
```{r}
b0b1_dist = res |> 
  ggplot(aes(x = log_beta01)) +
  geom_density() +
  labs(title = "Distribution of Bootstrapped log(beta0*beta1) Estimates", x = "log(beta0*beta1)", y = "Density")
b0b1_dist
```

This plot is also a density plot of bootstrapped Log Beta Product estimates. It follows the normal distribution.

## Problem 2
#### Import and Tidy Data
```{r}
birthweight = read_csv("data/birthweight.csv")
birth = 
  birthweight |> 
  janitor::clean_names() |> 
  mutate(babysex = case_when(babysex == 1 ~ 'male',
                             babysex == 2 ~ 'female'),
         frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
         malform = case_when(malform == 0 ~ 'absent',
                             malform == 1 ~ 'present'),
         mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other")) |> 
  na.omit()
```
## Regression Model
Assume all variables as predictors and perform MLR
```{r}
mlr = lm(bwt ~ ., data = birth)
mlr |> 
  broom::tidy() |> 
  knitr::kable()
```
#### Plot the model residuals against fitted values

```{r}
birth |> 
    add_predictions(mlr) |>  
    add_residuals(mlr) |> 
    ggplot(aes(x = pred, y = resid)) +
    geom_point()  + 
    geom_smooth(method = "lm") + 
    labs(title = "Residuals against fitted values", 
       x = "Fitted Values", 
       y = "Residuals")
```
We can find from the graph that the slope is 0, which means we have to choose predictors to measure the relationship.

#### Compare model to others
A model that employs only the baby's length at birth and gestational age as its main predictors, without considering any interaction effects between these variables.

A model that includes the baby's head circumference, length at birth, and sex as predictors, and additionally incorporates all possible interaction effects, including two-way interactions between each pair of these variables as well as the three-way interaction among head circumference, length, and sex.
```{r}
# model comparison 1
model_bg = lm(bwt ~ blength + gaweeks, data = birth) |> 
  broom::tidy() |> 
  knitr::kable()

# model comparison 2
model_bbb = lm(bwt ~ bhead * blength * babysex, data = birth) |> 
  broom::tidy() |> 
  knitr::kable()
model_bg
```
#### Compare these models by evaluating their cross-validated prediction errors. Utilize the `crossv_mc` function for cross-validation and apply relevant functions from the `purrr` package as needed for this analysis.
```{r}
cv_df = 
  crossv_mc(birth, 100) |> 
  mutate(
    mod1 = map(train, \(train) lm(bwt ~ blength + gaweeks, data = train) ),
    mod2 = map(train, \(train) lm(bwt ~ bhead * blength * babysex, data = train)),
    mod3 = map(train, \(train) lm(bwt ~ ., data = train))
  ) |> 
  mutate(
    rmse_mod1 = map2_dbl(mod1, test, \(mod, test) rmse(model = mod, data = test)),
    rmse_mod2 = map2_dbl(mod2, test, \(mod, test) rmse(model = mod, data = test)),
    rmse_mod3 = map2_dbl(mod3, test, \(mod, test) rmse(model = mod, data = test))
  )
cv_df
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Based on the analysis, it can be inferred that model3, which incorporates all predictors, exhibits the most favorable distribution of Root Mean Square Error (RMSE), showcasing the smallest RMSE values. Conversely, model1, which only uses birth length and gestational age as predictors, demonstrates the least favorable distribution, indicated by the largest RMSE values.






